---
title: "Week7_CompStats"
author: "Sharad Bajpai"
date: "11/2/2019"
output: 
  pdf_document: 
    latex_engine: xelatex
---

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```

1. Plot these four points using R.
```{r}
library(tidyverse)
Age<- c(23,18,10,45)
IQ<- c(100,105,95,120)
Table1 <- cbind(Age, IQ)
Table1<-as.data.frame(Table1)
view(Table1)
p <- ggplot(Table1, aes(x=Age, y=IQ)) + geom_point() +
  geom_smooth(method=lm, se=FALSE)+
  ggtitle("Age_IQ_Graph") +
  xlab("Age")+
  ylab("IQ")
p
```

2. Calculate the covariance between age and IQ.
$$\textrm{Cov}(x,y)=\frac{1}{1-n}\sum (x_i - \bar{x})(y_i - \bar{y})$$
$$Mean of ages = \frac{sumofages}{number} = \frac{23+18+10+45}{4}=24$$
$$Mean of IQ = \frac{sumofIQ}{number} = \frac{100+105+95+120}{4}=105$$
$$\textrm{Cov}(Age,IQ) = \frac{1}{(4-1) } [(23-24)(100-105)+(18-24)(105-105)+(10-24)(95-105)+(45-24)(120-105)]$$
$$\textrm{Cov}(Age,IQ) = \frac{460}{3} = 153.3333$$

Hence, COvariance between age and IQ is 153.3333

3. Calculate their correlation. What does the number you get indicate?

$$r = \frac{\textrm{Cov}(x,y)}{ s_{x} s_{y}}$$

S_x: Standard deviation of Age
S_y: Standard deviation of IQ

$$S_x=\sqrt{\frac{1}{n-1}\sum(x_i-\bar{x})^2}$$
$$S_x=\sqrt{\frac{1}{4-1}\sum(23-24)^2+(18-24)^2+(10-24)^2+(45-24)^2}$$
$$S_x=\sqrt{\frac{1}{3}*674}$$
$$S_x=14.9889$$
$$S_y=\sqrt{\frac{1}{n-1}\sum(y_i-\bar{y})^2}$$

$$S_y=\sqrt{\frac{1}{4-1}\sum(100-105)^2+(105-105)^2+(95-105)^2+(120-105)^2}$$
$$S_y=\sqrt{\frac{1}{3}*350}$$
$$s_y = 10.80123$$
Thus, Correlation is:

$$r = \frac{Cov_{(x,y)}}{s_x*s_y}=\frac{153.3333}{14.9889*10.80123}=0.9470956$$
Our, Correlation(r) = 0.9470956 value is very close to 1 which indicates that Age and IQ are strongly correlated.

4. Calculate the regression coefficients β0 and β1 and write out the equation of the best-fit line relating
age and IQ.

For $\beta_1$:

$$r = \frac{Cov_{(x,y)}}{s_x*s_y}=\beta_1\frac{s_x}{s_y}$$
$$\beta_1=r*\frac{s_y}{s_x}$$
$$\beta_1=\frac{0.9470956 * 10.80123}{14.9889}=0.682492$$

For $\beta_0$:

$$\beta_0=\bar{y}-\beta_1\bar{x}$$
$$\beta_0=105-(0.682492*24)=88.62019$$

Therefore, equation of the best-fit line relating age and IQ is:
$$\hat{y}_i = \beta_0 + \beta_1x_i$$
$$\hat{y}_i = 88.62019 + 0.682492 x_i$$


5. Calculate the predicted yˆi for each xi

For each $x_i$:
$$\hat{y}_i=\beta_0 + \beta_1x_i$$
$$\hat{y}_i= 88.62019 + (0.682492*23)=104.317506$$
$$\hat{y}_i= 88.62019 + (0.682492*18)=100.905046$$
$$\hat{y}_i= 88.62019 + (0.682492*10)=95.44511$$
$$\hat{y}_i= 88.62019 + (0.682492*45)=119.33233$$




6. Calculate Rˆ2 from the TSS/SSE equation. How does it relate to the correlation? What does the number you get indicate?

$$R^2 = \frac{TSS - SSE}{TSS} $$
Here I am calculating Total Sum of Squares:
$$TSS = \sum_{i} (y_{i} - \bar{y})^{2}=(100-105)^2+(105-105)^2+(95-105)^2+(120-105)^2=350$$
Here I am calculating Sum of Squared Errors:
$$SSE = \sum_{i} (y_{i} - \hat{y})^{2}=(100-104.317506)^2+(105-100.905046)^2+(95-95.44511)^2+(120-119.33233)^2 = 18.636 + 16.769025 + 0.19 + 0.44$$
$$SSE = 36.03$$

Therefore,
$$R^2 = \frac{TSS-SSE}{TSS} = \frac{350 - 36.03}{350}=0.8970$$

The $R^2$ value describes the proportion of the variation in the Y data that is explained by the values for X. It means how well the line of best fits predicts the data.

The $R^2$ value is also sometimes called the proportional reduction in error and is described as the proportional reduction of error in the variation in Y that would be explained by the line of best fit. A value of 0.897 suggests that ~89.7% of the variation in Y can be explained by X.



7. Calculate the standard error of β1, and use that to test (using the t test) whether β1 is significant.


$$se_{\hat{y}} = \sqrt{ \frac{\sum (y_i-\hat{y}_i)^2 }{n-2}}$$
$$se_{\hat{y}} = \sqrt{\frac{SSE}{n-2}}$$
$$ se_{\hat{y}} = \sqrt{ \frac{36.03}{4-2}} $$
$$e_{\hat{y}} = 4.24$$
$$ se_{\beta_0} = se_{\hat{y}} \sqrt{ \frac{\sum x_i^2}{n \sum (x_i - \bar{x})^2}}$$
$$se_{\beta_0} = 4.24 \sqrt{ \frac{2978}{4(674)}}$$
$$se_{\beta_1} = 4.24 \frac{1}{\sqrt{674}}$$
$$se_{\beta_1} = 0.16324$$ 
T-Test:
$$H_0:\beta_1=0$$ 
$$H_a:\beta_1\neq0$$  
$$t_{stat}=\frac{\beta_1 - \mu_{0}}{se_{\beta_1}}$$
$$t_{stat}=\frac{0.6824926 - 0}{0.16324}$$
$$t_{stat}=4.1809$$
$$df=n-k-1$$
$$df=4-1-1=2$$
$$t_{crit}=4.302653$$  
```{r}
qt(0.975,2)
```
We fail to reject the null hypothesis at the 95% confidence level with a 2-tailed test.

The positive correlation between age and IQ is not statistically significant. The failure to reject the null in this t-test of significance indicates that despite our Pearson coeffecient $r$ and proportional reduction in error $R^2$ suggesting that the linear regression model matches the data with reasonable accuracy, the positive correlation between age and IQ is not significant, and the positive correlation in this instance may be due to chance.


8. Calculate the p-value for β1 and interpret it.
```{r}
p_value<- 2*pt(4.18,2, lower.tail=F)
p_value
```
The $p_{value}=0.0527458$ suggests that if sample data was gathered on age and IQ with $n>100$ , about ~5.2% of those trials would have a mean less than or equal to the one in this trial. Since $\alpha=.05$, we conclude that at the 95% confidence level $\alpha/p=.05$ with a 2-tailed test, the result for this trial is not statistically significant.


9. Calculate the 95% CI for β1 and interpret it.

For 95% CI
$$CI_{95}=\beta_1\pm(ThresholdValue)*se_{\beta_1} $$
$$CI_{95}=0.682492 \pm (4.302653*0.16324)$$
$$CI_{95}=(-0.19873, 1.38485)$$
The 95% confidence interval for $\beta_1$ indicates that for each year of age we would expect that IQ would increase between $(-0.19873, 1.38485)$.
One end is very close to 0, suggesting that as each year passes, IQ may not increase hardly at all, and the other end suggesting a gain of up to 1.36 IQ points per year.



10. Confirm your results by regressing IQ on Age using R.
```{r}
summary(lm(IQ~Age,data=Table1))
```

Confirmation: The results of the lm function matches the previous calculation.
The P-value is greater than alpha=0.05 so we cannot reject the null hypothesis.

11. Plot your points again using R, including the linear fit line 
with its standard error.
```{r}
ggplot(Table1, aes(x= Age, y= IQ))+ geom_point()+ geom_smooth(method=lm)+ggtitle("Graph of Age and IQ")+ theme(plot.title = element_text(hjust=0.5))
```
The shaded area in the above graph represents a measure in variation in linear fit line with its standard error.

12. What are you final conclusions about the relationship between 
age and IQ?
After the analysis, the final conclusion about the relationship between Age and IQ is :

$r = 0.947$ indicating there is a strong positive correlation between age and IQ.
$R^2=0.897$ indicates that approximately 89.7% of the error has been reduced and is explained by the line of best fit.
The t-test at the 95% confidence level yielding a p-value of $~.0527$ indicates that despite the strength of the correlation indicated by the $r$ and $R^2$ values, this correlation is not signficant. It may not actually be as positive as the line of best fit in this trial suggests.

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```
